---
title: "Project 1 - Chess Player DB"
author: "Sedo Oh"
date: "September 21, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read in Text file
Being very sick this week, I took the most simplest approach.

First read in the txt file using | as delimiter.  
In addition, align the data so that all the information needed is in a single row instead of two.
Also drop all the columns that's not needed.

```{r Read in Data}
tournament_raw <- read.delim("C:\\Users\\soh1\\Box Sync\\CUNY\\Fall 2017\\Data 607\\Week3\\Project\\tournamentinfo.txt", sep="|", header = FALSE, stringsAsFactors = FALSE)

tournament_raw <- tournament_raw[5:nrow(tournament_raw),]

name_row <- c()
for (i in seq(1,nrow(tournament_raw), by=3)){
  name_row <- rbind(name_row,tournament_raw[i,]) 
}

state_row <- c()
for (i in seq(2,nrow(tournament_raw), by=3)){
  state_row <- rbind(state_row,tournament_raw[i,]) 
}

name_state <- cbind(name_row, state_row)

##clean up
name_state <- name_state[,1:13]
names(name_state) <- c("playerid","name", "Total Pts", "Round1", "Round2", "Round3", "Round4","Round5","Round6","Round7","drop","state","rank")
name_state <- name_state[,!(names(name_state) %in% "drop")]

head(name_state)
```

## Extract Rating and Opponent Players

I called Rating, ranking for now.  Using str_split_fixed to separate out Rating string and keep only the Pre-Rate. 
Using str_extract to get the opponent playerid (keep numbers only)
```{r Extract Rating and Opponent Players}
##extract rank
library(stringr)
library(data.table)
rank <- data.table((str_split_fixed(name_state$rank," ",5)))
rank_final <- data.table(str_extract(rank$V5,"[:digit:]+"))
name_state$rank <- rank_final$V1

##extract players played against
playerlist <- c()
for (i in 1:7){
  Roundname <- name_state[,names(name_state) %in% paste0("Round",i)]
  Roundname <- str_extract(Roundname,"[:digit:]+")
  name_state[,names(name_state) %in% paste0("Round",i)] <- Roundname
}

head(name_state)
```


## Merge Rating of the Opponent Players
First, format the db so that the variables can be used for calculation and lookup.

Create another table with just playerid and rating.  Then just merge on the rating by each Round's opponent players.

```{r Merge ranking of the opponent players}
##Format 
name_state$playerid <- as.integer(name_state$playerid)
name_state[,4:10] <- lapply(name_state[,4:10], as.integer)
name_state$rank <- as.integer(name_state$rank)

##Another table of playerid and rank and merge
ranking_table <- name_state[,c("playerid","rank")]

##merge
for (i in 1:7){
  names(ranking_table) <- c(paste0("Round",i), paste0("rank",i))
  name_state <- merge(name_state,ranking_table, paste0("Round",i), all.x = TRUE)
}

head(name_state)
```

## Calculate the Mean and create Final DB

Here I renamed some variables to be consistent with definition of the text file.
```{r mean and final db}
##calculate the mean
name_state$avgOppo <- rowMeans(name_state[,13:19], na.rm=TRUE)

##Final condense DB
FinalDB <- name_state[,c("name","state","Total Pts", "rank","avgOppo")]
names(FinalDB) <- c("name","state","Total Pts", "Pre_Rating","Avg Pre Opponent Rating")

head(FinalDB)
```

