---
title: "TidyPractice"
author: "Sedo Oh"
date: "September 23, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Tidy Practice

Read in the data

```{r cars}
setwd("C:\\Users\\soh1\\Box Sync\\CUNY\\Fall_2017\\Data 607\\Week5")
AirlineRaw <- read.csv("AirlineTime.csv")
```

Gather City into one column. And then spread out delay vs on time on separate columns

```{r restructure}
library(tidyr)
library(dplyr)
library(tidyselect)

AirlineTrans <- AirlineRaw %>%
  gather(City, Frequency, 3:7) %>%
  spread(Time, Frequency)
```

Let's also summarize by Airlines
```{r summarize}
AirlineSum <- AirlineTrans %>%
  group_by(Airplane) %>%
  summarise(TotalDelay = sum(delayed))
```

Now let's use ggplot to compare the airlines performance
```{r Airline Overall Performance}
library(ggplot2)
ggplot(AirlineSum) + aes(Airplane,TotalDelay) + geom_bar(stat="identity", position=position_dodge())

```

AM West is doing much worse than Alaska overall.

Let's take a look at how airlines perform by the city.
```{r Airline by City}
ggplot(AirlineTrans) + aes(City,delayed, fill=Airplane) + geom_bar(stat="identity", position=position_dodge())

```

AM West does terribly in Phoenix and Alaska does terribly in Seattle.  
